### 2020-05-25 writeup/transfer.tex
sec 2 wip

We illustrate our main results (to be presented in Section \ref{sec_main}) by considering a few special cases, namely special settings of the task models $\set{\beta_i}_{i=1}^k$, covariance matrices $\set{\Sigma_i}_{i=1}^k$, and number of data points $\set{n_i}_{i=1}^k$.
We show that our results explain several phenomenon that cannot be explained before.
\todo{list those here}

@@

We assume that for every row $x^\top$ of $X_i$, we have $\ex{xx^{\top}} = \Sigma_i.$
We also write $x = \Sigma_i^{1/2} z_i$, where $z_i$ is a random vector that has i.i.d. entries with mean $0$ and variance $1$.
We will designate the $k$-th task as the target.
Our goal is to come up with an estimator $\hat{\beta}$ to provide accurate predictions for the target task, provided with the other auxiliary task data.
Concretely, we focus on the test error for the target task:
	\begin{align*}
		\te_k(\hat{\beta}) &\define \exarg{x \sim \Sigma_k}{\exarg{{\varepsilon_i, \forall 1\le i\le k}}{(x^{\top}\hat{\beta} - x^{\top}\beta_t)^2}} \\
		&= \exarg{\varepsilon_i, \forall 1\le i\le k}{(\hat{\beta} - \beta_t)^{\top}\Sigma_k(\hat{\beta} - \beta_t)}.
	\end{align*}
%\end{itemize}

@@

\textbf{The High-Dimensional Setting.}
%\todo{setup motivation and notations}
We would like to get insight on how covariate and model shifts affect the rate of transfer.
We will consider the high-dimensional setting where for the target task, its number of data points is a small constant times $p$.
This setting captures a wide range of applications of multi-task learning where we would like to use auxiliary task data to help train tasks with limited labeled data.
Furthermore, this setting is particularly suited to our study since there is need for adding more data to help learn the target task.

@@

\textbf{Two Tasks with General Covariance Matrices.}
For the case of two tasks, we decompose the test error of $\hat{\beta}_{t}^{\MTL}$ on the targe task into two parts
\begin{align}
	\te(\hat{\beta}_t^{\MTL}) =& \hat{v}^2 \bignorm{\Sigma_2^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1 (\beta_s - \hat{v}\beta_t)}^2 \nonumber \\
	&+ \sigma^2\cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}, \label{eq_te_model_shift}
\end{align}
where $\hat{v}$ denotes the ratio of the output layer weights (to be defined more precisely in Appendix \ref{app_proof_main}).
It is not hard to see that the variance of $\hat{\beta}_t^{\MTL}$ is reduced compared to $\hat{\beta}_t^{\STL}$, i.e. %(following the argument of Proposition \ref{prop_monotone}), i.e.
\[ \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} \le \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2}. \]
Because of model shift bias, we can no longer guarantee that $\te(\hat{\beta}_{t}^{\MTL}) \le \te(\hat{\beta}_t^{\STL})$.
%The main result of this section show deterministic conditions under which we get positive or negative transfer.
%And the conditions depend only on the covariate shift matrix $M$, the difference of the task models, and the number of per-task data points.
The technical crux of our approach is to derive the asymptotic limit of $\te(\hat{\beta}_t^{\MTL})$ in the high-dimensional setting, when $p$ approaches infinity.
A key highlight of our approach implies a precise limit on $\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}}$, which only depends on $\Sigma_1, \Sigma_2$ and $n_1, n_2$ (see Lemma \ref{lem_cov_shift} in Appendix \ref{app_proof_main} for the result).

@@

\textbf{Many Tasks with the Same Covariates.}
We extend the above result to any number of tasks that have the same covariates.
%In this section we consider the setting with $k$ many that have the same covariates.
Since the tasks all have the same number of datapoints and covariance matrix, the trade-off between model shift bias and variance will be captured by their task models $\set{\beta_i}_{i=1}^k$.
%For this setting, we derive solutions for the multi-task training and the transfer learning setting that match our insights qualitatively from Section \ref{sec_denoise}.
Let $B^{\star} = [\beta_1, \beta_2, \dots, \beta_k] \in\real^{p\times k}$ denote the underlying task model parameters.
We derive the model shift bias and variance in the following result.

>>>>

We revisit the classical bias-variance trade-off in the context of multi-task learning for linear regressions, in particular the setup of equation \eqref{eq_mtl}.
For the high-dimensional setting where the data size of the target task is a small constant times dimension $p$, we quantify the trade-off as a function of the task models $\set{\beta_i}_{i=1}^k$, covariance matrices $\set{\Sigma_i}_{i=1}^k$, and number of data points $\set{n_i}_{i=1}^k$.
After describing this technical result, we provide theoretical insights over when multi-task learning performs better than single-task learning.
Our analysis reveals three contributing causes of negative transfer, including task model \textit{dissimilarity}, \textit{unbalanced} data size, and \textit{covariate shift}, which we describe in detail for the rest of the section.

@@

\textbf{Notations.}
We assume that for every row $x^\top$ of $X_i$, we have $\ex{xx^{\top}} = \Sigma_i.$
We also write $x = \Sigma_i^{1/2} z_i$, where $z_i$ is a random vector that has i.i.d. entries with mean $0$ and variance $1$.
We will designate the $k$-th task as the target.
Our goal is to come up with an estimator $\hat{\beta}$ to provide accurate predictions for the target task, provided with the other auxiliary task data.
Concretely, we focus on the test error for the target task:
	\begin{align*}
		\te_k(\hat{\beta}) &\define \exarg{x \sim \Sigma_k}{\exarg{{\varepsilon_i, \forall 1\le i\le k}}{(x^{\top}\hat{\beta} - x^{\top}\beta_t)^2}} \\
		&= \exarg{\varepsilon_i, \forall 1\le i\le k}{(\hat{\beta} - \beta_t)^{\top}\Sigma_k(\hat{\beta} - \beta_t)}.
	\end{align*}
%\end{itemize}

@@

%\textbf{The High-Dimensional Setting.}
%\todo{setup motivation and notations}
We would like to get insight on how covariate and model shifts affect the rate of transfer.
We will consider the high-dimensional setting where for the target task, its number of data points is a small constant times $p$.
This setting captures a wide range of applications of multi-task learning where we would like to use auxiliary task data to help train tasks with limited labeled data.
Furthermore, this setting is particularly suited to our study since there is need for adding more data to help learn the target task.

@@

\textbf{Setting one: Two tasks with general covariance matrices.}
For the case of two tasks, we decompose the test error of $\hat{\beta}_{t}^{\MTL}$ on the targe task into two parts
\begin{align}
	\te(\hat{\beta}_t^{\MTL}) =& \hat{v}^2 \bignorm{\Sigma_2^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1 (\beta_s - \hat{v}\beta_t)}^2 \nonumber \\
	&+ \sigma^2\cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}, \label{eq_te_model_shift}
\end{align}
where $\hat{v}$ denotes the ratio of the output layer weights (to be defined more precisely in Appendix \ref{app_proof_main}).
It is not hard to see that the variance of $\hat{\beta}_t^{\MTL}$ is reduced compared to $\hat{\beta}_t^{\STL}$, i.e. %(following the argument of Proposition \ref{prop_monotone}), i.e.
\[ \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} \le \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2}. \]
Because of model shift bias, we can no longer guarantee that $\te(\hat{\beta}_{t}^{\MTL}) \le \te(\hat{\beta}_t^{\STL})$.
%The main result of this section show deterministic conditions under which we get positive or negative transfer.
%And the conditions depend only on the covariate shift matrix $M$, the difference of the task models, and the number of per-task data points.
The technical crux of our approach is to derive the asymptotic limit of $\te(\hat{\beta}_t^{\MTL})$ in the high-dimensional setting, when $p$ approaches infinity.
A key highlight of our approach implies a precise limit on $\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}}$, which only depends on $\Sigma_1, \Sigma_2$ and $n_1, n_2$ (see Lemma \ref{lem_cov_shift} in Appendix \ref{app_proof_main} for the result).

@@

\textbf{Setting two: Many tasks with the same covariates.}
We extend the above result to any number of tasks that have the same covariates.
%In this section we consider the setting with $k$ many that have the same covariates.
Since the tasks all have the same number of datapoints and covariance matrix, the trade-off between model shift bias and variance will be captured by their task models $\set{\beta_i}_{i=1}^k$.
%For this setting, we derive solutions for the multi-task training and the transfer learning setting that match our insights qualitatively from Section \ref{sec_denoise}.
Let $B^{\star} = [\beta_1, \beta_2, \dots, \beta_k] \in\real^{p\times k}$ denote the underlying task model parameters.
We derive the model shift bias and variance in the following result.
